---
title: "Turnover of species and interactions"
author: "Fernando Cagua"
date: "10/23/2018"
output:
  md_document:
    variant: markdown_github
---

# Interactions at the fringes, data exploration

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, dpi = 300, fig.asp = 10/16)
```

We want to investigate whether species interact differently at the core of their range compared to the fringes of their distribution. 
One could expect species to diversify their interactions as interaction parters become less common/available. Alternatively, one could expect species to interact with only a subset of the interaction partners, as they cannot develop new interactions when environemntal conditions are close to the limits of where possitive groth rates occur. 

The aim of this document is to explore whether the interaction data available in the Bascompte's Web of Life could be adequate to answer this question or not. 

## Species & interaction turnover

We start by checking wether species and interactions occur in enough localities. 
Unknown and unindentified species are excluded from the final numbers. 

If species turnover is too high the data might not be adequate to answer the research question. 
At first we set an arbitrary, but reasonable threshold of **five** locations. 
We need species to be present in at least this number of locations to have a relatively OK coverage across their range. 

With this threshold of five we have 13 plants and 41 pollination species that could be candidates to be focal species. 

```{r}
require(ggplot2)
drake::loadd(species_list)

th <- 5

plot_freq_step <- function(freq_df, th = 5){
  if ("guild" %in% names(freq_df)){
    p <- freq_df %>%
      ggplot(aes(x = n_loc_rank, y = n_loc, colour = guild)) 
  } else {
    p <- freq_df %>%
      ggplot(aes(x = n_loc_rank, y = n_loc)) 
  }
  p +
    geom_step(size = 1) +
    scale_x_log10() +
    theme_bw() +
    scale_color_brewer(palette = "Paired", 
                       aesthetics = c("colour", "fill")) +
    scale_alpha_manual(values = c(0.5, 1)) +
    scale_size_manual(values = c(0.5, 1)) +
    labs(x = "rank", 
         y = "# locations") +
    theme(legend.position = c(0.98,0.98), 
          legend.justification = c(1,1), 
          legend.direction = "horizontal")
}

sp_rank <- species_list %>%
  remove_unknown_species() %>% 
  get_sp_freq_by_group(sp_name)  %>%
  dplyr::mutate(focal = n_loc >= th) 

n_sp <- sp_rank %>%
  dplyr::group_by(guild, focal) %>%
  dplyr::summarise(n_sp = dplyr::n_distinct(sp_name)) %>%
  dplyr::mutate(label = paste(format(n_sp, big.mark = ","), guild)) %>%
  dplyr::group_by() %>%
  humanize()

sp_rank %>%
  dplyr::group_by() %>%
  humanize() %>%
  plot_freq_step() +
  geom_vline(data = n_sp, aes(xintercept = n_sp, 
                              linetype = focal), 
             size = 0.25, colour = "grey50") +
  geom_text(data = n_sp, 
            aes(label = label, x = n_sp, y = 15),
            size = 3.5,
            angle = 90, 
            nudge_x = -0.05, show.legend = FALSE) +
  geom_text(data = n_sp,
            aes(label = focal, x = n_sp, y = 15), 
            angle = 90, nudge_x = 0.05, colour = "grey30", 
            size = 3) +
  scale_linetype_manual(values = c(3,2)) +
  guides(size = "none", linetype = "none") +
  labs(title = "number of locations per species", 
       subtitle = "species turnover in the Web of Life database", 
       caption = "we select as focal species those in 5 or more locations")

```

We then check the turnover of the interactions. 
We expect a large degree of turnover here, but if we are to compare interactions at the fringes of a species distributions with those at the core, we need at some of the interactions to be sampled at several locations. These  would compose the "core" of the species interactions.
At first we set an arbitrary, but reasonable threshold of **three** locations.

With this threshold we have 44 interactions that repeat over at least three locations or more. 
Specifically 22 plants and 31 pollinators are involved in these interactions. 

```{r}
require(ggplot2)
drake::loadd(interaction_list)

int_rank <- interaction_list %>%
  remove_unknown_species() %>% 
  get_int_freq_by_group(pla_name, pol_name) %>%
  dplyr::mutate(focal = n_loc >= 3)

n_int <- int_rank %>%
  dplyr::group_by(focal) %>%
  dplyr::summarise(n_int = dplyr::n_distinct(paste(pla_name, pol_name))) %>%
  dplyr::mutate(label = paste(format(n_int, big.mark = ","), "int."))

n_int_sp <- int_rank %>%
  dplyr::filter(focal) %>%
  dplyr::select_at(dplyr::vars(dplyr::contains("name"))) %>%
  tidyr::gather("guild", "sp_name") %>%
  dplyr::mutate(guild = substr(guild, 1,3)) %>%
  dplyr::group_by(guild) %>%
  dplyr::summarise(n_sp = dplyr::n_distinct(sp_name)) %>% {
    l <- .
    as.list(extract2(l, 2)) %>%
      `names<-`(extract2(l, 1))
  }

int_rank %>%
  plot_freq_step() +
  # geom_hline(yintercept = 2.9, size = 0.5, linetype = 2) +
  geom_vline(data = n_int, aes(xintercept = n_int, 
                                linetype = focal), 
             size = 0.25, colour = "grey50", show.legend = F) +
    scale_linetype_manual(values = c(3,2)) +
  annotate(geom = "text", x = 46, y = 3.5, label = paste(n_int_sp[[1]], "plants and", n_int_sp[[2]], "pollinators with the\nsame int. in more than 3 locations"), size = 3, hjust = 0, colour = "grey30") +
  labs(title = "locations per interactions", 
       subtitle = "interactions turnover in the Web of Life database", 
       caption = "we select as focal species those with the same interaction in 3 or more locations")
```

Combining these two thresholds

 * Only species with interactions that were sampled in three locations or more
 * Only species sampled in 5 locations or more

we have 16 focal species for which enough interaction data could be obtained from Bascompte's Web of Life. Four plants and 12 pollinators. 

Now let's have a little look at how these numbers depend on the chosen threshold. 

```{r}
drake::loadd(focal_species_options)

focal_species_options %>%
  dplyr::group_by(int_threshold, sp_threshold, guild) %>%
  dplyr::summarise(n = n()) %>% 
  dplyr::group_by(int_threshold, sp_threshold) %>%
  dplyr::mutate(n_total = sum(n), 
                label = paste(format(n_total, big.mark = ","), "spp."),
                label2 = paste(format(n, big.mark = ","), guild), 
                label2 = paste(label2, collapse = "+")) %>%
  dplyr::select(-guild) %>%
  dplyr::distinct() %>%
  humanize() %>%
  ggplot(aes(x = int_threshold, y = sp_threshold)) +
  geom_tile(aes(fill = n_total)) +
  geom_text(aes(label = label), 
            size = 3, 
            fontface = "bold", 
            nudge_y = 0.2, 
            color = "white") +
  geom_text(aes(label = label2), 
            size = 3, 
            nudge_y = -0.2, 
            color = "white") +
  geom_tile(aes(x = 3, y = 5), fill = NA, colour = "white", linetype = 2) +
  scale_fill_continuous(trans = "log10") +
  coord_cartesian(ylim = c(1,10)) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(breaks = 1:10) +
  theme_bw() +
  labs(title = "number of focal species", 
       subtitle = "number of plants and pollinators using different thresholds", 
       x = "repeated interaction threshold", 
       y = "species location threshold", 
       fill = "# spp.", 
       caption = "dashed box indicate thresholds used here")
```

## Degree vs. widespread

Let's have a closer look at the chosen thresholds of 5 and 3. 
I'm a bit worried that we're only selecting super generalist species and hence results might be biased. 
We can see that indeed the overall species degree of the focal species is considerably larger than that of non-focal. We have still a decent range between 6 for *Fuchia regia* and like 400 for *Apis melifera*. 

```{r}
degrees <- interaction_list %>%
  dplyr::group_by(pla_name) %>%
  dplyr::mutate(pla_deg = dplyr::n_distinct(pol_name)) %>%
  dplyr::group_by(pol_name) %>%
  dplyr::mutate(pol_deg = dplyr::n_distinct(pla_name))%>%
  dplyr::group_by()

guilds <- c("pla", "pol")

sp_degree <- guilds %>%
  purrr::map(~dplyr::select_at(degrees, dplyr::vars(dplyr::starts_with(.)))) %>%
  purrr::map(~dplyr::distinct(.)) %>%
  purrr::map_df(~`names<-`(., c("sp_name", "genus", "deg")), .id = "guild") %>%
  dplyr::mutate(guild = factor(guild, labels = guilds), 
                guild = as.character(guild)) %>%
  remove_unknown_species()

threshold <- 3
# select species with interactions that occur in 3 locations or more
sp_int_thres <- int_rank %>%
  dplyr::filter(n_loc >= threshold) %>%
  dplyr::select_at(dplyr::vars(dplyr::ends_with("name"))) %>%
  tidyr::gather("guild", "sp_name", pla_name, pol_name) %>%
  dplyr::mutate(guild = substr(guild, 1,3)) %>%
  dplyr::distinct() 

# select species that occur in 3 locations or more
sp_thres <- sp_rank %>%
  dplyr::filter(n_loc >= 5) %>%
  dplyr::select(guild, sp_name)

selected_species <- dplyr::inner_join(sp_int_thres, sp_thres) %>%
  dplyr::mutate(selected = TRUE)

sp_deg_rank <- sp_rank %>%
  dplyr::inner_join(sp_degree) %>% 
  dplyr::left_join(selected_species) %>%
  dplyr::mutate(selected = dplyr::if_else(is.na(selected), FALSE, selected)) 
```

```{r}
require(ggplot2)
require(ggrepel)

median_deg <- sp_deg_rank %>%
  dplyr::group_by(selected, guild) %>%
  dplyr::summarise(deg = median(deg)) %>%
  dplyr::mutate(focal = selected) %>%
  dplyr::group_by() %>%
  humanize()

sp_deg_rank %>%
  dplyr::group_by() %>%
  humanize() %>%
  ggplot(aes(x = deg, y = n_loc)) +
  geom_vline(data = median_deg, aes(xintercept = deg, colour = selected), 
             size = 0.25, 
             linetype = 2) +
  geom_text(data = median_deg, 
            aes(y = 20, 
                label = paste("median", guild, "-", deg), 
                color = selected),
            nudge_x = -0.05,
            angle = 90, 
            size = 3) +
  geom_point(aes(colour = selected, fill = selected, shape = guild)) +
  geom_label_repel(data = dplyr::filter(sp_deg_rank, selected), 
                  aes(label = sp_name), size = 3, fontface =  "italic",
                  color = RColorBrewer::brewer.pal(3, "Paired")[2]) +
  scale_x_log10() +
  # scale_y_log10() +
  scale_fill_manual(values = c(NA, RColorBrewer::brewer.pal(3, "Paired")[2])) +
  scale_color_brewer(palette = "Paired") +
  scale_shape_manual(values = c(21, 24)) +
  guides(color = guide_legend("focal species"), 
         fill = guide_legend("focal species")) +
  theme_bw() +
  theme(legend.position = c(0.02,0.98), 
        legend.justification = c(0,1), 
        legend.box = "horizontal", 
        legend.direction = "horizontal") +
  labs(title = "species degree vs. number of locations", 
       subtitle = "there are 16 focal species, mostly generalist pollinators",
       caption = "focal species are those that have the same interaction in >= 3 locations and are present >= 5 locations", 
       y = "# locations", 
       x = "species degree")
```

## Data completeness

```{r}
drake::loadd(n_occurrences)

n_occurrences %>%
  dplyr::group_by(sp_name, n, guild, n_occurrences) %>%
  dplyr::do(canon_name = unlist(.$backbones[[1]]$canonicalName)) %>% 
  dplyr::group_by() %>%
  dplyr::mutate(sp_name = forcats::fct_reorder(sp_name, n), 
                n_occurrences = dplyr::if_else(sp_name != canon_name, 
                                               NA_integer_, 
                                               as.integer(n_occurrences))) %>%
  humanize() %>%
  ggplot(aes(y = n_occurrences, x = sp_name, colour = guild)) +
  geom_point() +
  geom_segment(aes(xend = sp_name, yend = 0)) +
  scale_y_log10() +
  coord_flip() +
  scale_color_brewer(palette = "Paired") +
  theme_bw() +
  theme(axis.text.y = element_text(face = "italic", size = 6)) +
  labs(x = "", y = "# occurrences", 
       title = "GBIF occurrences", 
       subtitle = "number of GBIF georeferenced records per species", 
       caption = "species ordered by how likely they are to be included as focal species")
```

