---
title: "Flexibility of interactions depends on habitat suitability"
author: 
  - "E. Fernando Cagua^1^ ([efc29@uclive.ac.nz](mailto:efc29@uclive.ac.nz))"
  - "Audrey Lustig^2^ ([audrey.lustig@canterbury.ac.nz](audrey.lustig@canterbury.ac.nz))"
  - "Jason M. Tylianakis^1^ ([jason.tylianakis@canterbury.ac.nz](jason.tylianakis@canterbury.ac.nz))"
  - "Daniel B. Stouffer^1^ ([daniel.stouffer@canterbury.ac.nz](daniel.stouffer@canterbury.ac.nz))"
bibliography: 
  - bibliography.bib
  - int-bibliography.bib
csl: pnas.csl
output: #word_document
  bookdown::pdf_document2:
    # base_format: rticles::peerj_article
    keep_tex: yes
    number_sections: false
toc: false
fontsize: 11pt
# classoption: a4paper
geometry: 
  - textwidth=33em
  - textheight=48\baselineskip
header-includes:
  - \usepackage{booktabs}
  - \usepackage{setspace}
  - \usepackage{lineno}
  - \usepackage{xr}
  - \usepackage[utf8]{inputenc}
  - \newcommand{\R}[1]{\label{#1}\linelabel{#1}}
  - \newcommand{\lr}[1]{page~\pageref{#1}, line~\lineref{#1}}
  - \externaldocument[S-]{supp-info}
  - \usepackage[export]{adjustbox}
---

\linenumbers
\doublespacing

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, 
                      warning = FALSE, dpi = 300, 
                      eval.after = "fig.cap")

loadd(abs_wordcount, msc_wordcount, n_displays, n_references, keywords)
loadd(interaction_citations)
```
 
^1^ Centre for Integrative Ecology, School of Biological Sciences, University of Canterbury, Private Bag 4800, Christchurch 8041, New Zealand

^2^ Geospatial Research Institute, University of Canterbury, Private Bag 4800, Christchurch 8041, New Zealand

**Running title:** XX

**Keywords:** `r keywords`

**Type of article:** XX

**Number of words:** `r abs_wordcount$n_words_stri` in abstract; `r format(round(msc_wordcount$n_words_stri), big.mark = ",")` in main text.

**Number of displays:** `r n_displays$figures` figures; `r n_displays$tables` tables; `r n_displays$text_boxes` text boxes.

**Number of references:** `r n_references + nrow(interaction_citations)`

**Author for correspondence:** E. Fernando Cagua (+64 20 4026 8153).

**Data accessibility:** Data supporting the results will be accessible in an appropriate data repository after publication. The DOI will be included here.

**Author contributions:** XX

\clearpage

# Abstract

`r paste(drake::readd(abstract), collapse = "")`

\clearpage

# Introduction

# Methods

```{r}
loadd(config)
loadd(int_metadata)

n_published_studies <- int_metadata$reference %>%
  unique() %>%
  stringr::str_detect(pattern = "unpubl", negate = TRUE) %>%
  sum()

```

We retrieved plant-pollinator networks from the Web of Life database [@fortuna_web_2014] on`r format(as.Date(config$raw_data_retrieved), "%b %d, %Y")` (Figure \@ref(fig:fig-worldmap)). 
Some studies included interaction data recorded at multiple points in time. 
When this was the case, we combined them into one single interaction network per location. 
All together, these networks included 

```{r fig-worldmap, fig.width = width('single'), fig.cap = paste("Worldwide distribution of pollination communities included in this study. We compile data from", length(interaction_citations$BIBTEXKEY), "published studies", all_int_cites, "which contain interaction information between plant and their pollinators at", dplyr::n_distinct(int_metadata$loc_id), "distinct locations.")}
readd(fig_worldmap)

all_int_cites <- paste0("@", interaction_citations$BIBTEXKEY) %>%
  glue::glue_collapse(sep = "; ") %>%
  paste0("[", ., "]")

```

Studies are 

# Results

# Discussion 

# Acknowledgements

`r paste(drake::readd(acknowledgements), collapse = "")`

# References
