---
title: "Flexibility of interactions depends on habitat suitability"
author: 
  - "E. Fernando Cagua^1^ ([efc29@uclive.ac.nz](mailto:efc29@uclive.ac.nz))"
  - "Audrey Lustig^2^ ([audrey.lustig@canterbury.ac.nz](audrey.lustig@canterbury.ac.nz))"
  - "Jason M. Tylianakis^1^ ([jason.tylianakis@canterbury.ac.nz](jason.tylianakis@canterbury.ac.nz))"
  - "Daniel B. Stouffer^1^ ([daniel.stouffer@canterbury.ac.nz](daniel.stouffer@canterbury.ac.nz))"
bibliography: 
  - bibliography.bib
  - int-bibliography.bib
csl: pnas.csl
output: #word_document
  bookdown::pdf_document2:
    # base_format: rticles::peerj_article
    keep_tex: yes
    number_sections: false
toc: false
fontsize: 11pt
# classoption: a4paper
geometry: 
  - textwidth=33em
  - textheight=48\baselineskip
header-includes:
  - \usepackage{booktabs}
  - \usepackage{setspace}
  - \usepackage{lineno}
  - \usepackage{xr}
  - \usepackage[utf8]{inputenc}
  - \newcommand{\R}[1]{\label{#1}\linelabel{#1}}
  - \newcommand{\lr}[1]{page~\pageref{#1}, line~\lineref{#1}}
  - \externaldocument{supp-info}
  - \usepackage[export]{adjustbox}
editor_options: 
  chunk_output_type: console
---

\linenumbers
\doublespacing

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, 
                      warning = FALSE, dpi = 300, 
                      eval.after = "fig.cap")

library(dplyr)

loadd(abs_wordcount, msc_wordcount, n_displays, n_references, keywords)
loadd(interaction_citations)
```
 
^1^ Centre for Integrative Ecology, School of Biological Sciences, University of Canterbury, Private Bag 4800, Christchurch 8041, New Zealand

^2^ Geospatial Research Institute, University of Canterbury, Private Bag 4800, Christchurch 8041, New Zealand

**Running title:** XX

**Keywords:** `r keywords`

**Type of article:** XX

**Number of words:** `r abs_wordcount$n_words_stri` in abstract; `r format(round(msc_wordcount$n_words_stri), big.mark = ",")` in main text.

**Number of displays:** `r n_displays$figures` figures; `r n_displays$tables` tables; `r n_displays$text_boxes` text boxes.

**Number of references:** `r n_references + nrow(interaction_citations)`

**Author for correspondence:** E. Fernando Cagua (+64 20 4026 8153).

**Data accessibility:** Data supporting the results will be accessible in an appropriate data repository after publication. The DOI will be included here.

**Author contributions:** XX

\clearpage

# Abstract

`r paste(drake::readd(abstract), collapse = "")`

\clearpage

# Introduction

# Methods

## Data aquisition

### Species interactions

```{r}
loadd(config)
loadd(int_metadata)

n_published_studies <- int_metadata$reference %>%
  unique() %>%
  stringr::str_detect(pattern = "unpubl", negate = TRUE) %>%
  sum()

all_int_cites <- paste0("@", interaction_citations$BIBTEXKEY) %>%
  glue::glue_collapse(sep = "; ") %>%
  paste0("[", ., "]")

int_citations_year_range <- range(interaction_citations$YEAR)
```

```{r scientific-names-cleaning-info}
loadd(checked_sp_names, spp, manual_name_corrections, problematic_networks)

n_sp_names <- n_distinct(spp$sp_name)

n_sp_names_sp_level <- filter(spp, !sp_unidentified) %$% 
  n_distinct(sp_name)

n_sp_names_gen_level <- filter(spp, sp_unidentified, !gen_unidentified) %$% 
  n_distinct(sp_name)

n_no_tax_info <- filter(spp, gen_unidentified) %$% n_distinct(sp_name) 

n_sp_correct_in_data <- checked_sp_names %>%
  filter(good_queried_sp_name) %$%
  n_distinct(queried_sp_name)

n_sp_corrected_typos <- checked_sp_names %>%
  filter(corrected_typo) %$% 
  n_distinct(queried_sp_name)

sp_with_synonyms <- checked_sp_names %>%
  filter(alt_sp_name_found) 

n_sp_with_synonyms <- sp_with_synonyms %$%
  n_distinct(queried_sp_name)

sp_not_found <- checked_sp_names %>%
  filter(!good_queried_sp_name) %>% 
  filter(!alt_sp_name_found, !corrected_typo) 

n_sp_not_found <- sp_not_found %$%
  n_distinct(queried_sp_name)

n_sp_manual_corrections <- manual_name_corrections %>%
  filter(!is.na(sp_name)) %$%
  n_distinct(queried_sp_name)

study_per_location <- int_metadata %>%
  distinct(loc_id, reference)

n_names_in_multiple_references_orig <- spp %>%
  left_join(study_per_location, by = "loc_id") %>%
  group_by(sp_name) %>%
  summarise(n_net = n_distinct(loc_id), 
            n_ref = n_distinct(reference)) %>%
  filter(n_ref > 1) %$%
  n_distinct(sp_name)
```

We retrieved plant-pollinator networks from the Web of Life database [@fortuna_web_2014]. 
This database contains datasets originating from `r length(interaction_citations$BIBTEXKEY)` studies published in the primary literature between `r int_citations_year_range[1]` and `r int_citations_year_range[2]` `r all_int_cites`. 
All together, these data sources contained `r format(n_sp_names, big.mark = ",")` unique organism names; `r format(n_names_in_multiple_references_orig, big.mark = ",")` of them in more than one data source.  
Because assessing how habitat suitability affects interaction partners of plants and pollinators requires linking of species across different datasets, it was important to ensure that the cannnonical names used to identify species were as accurate as possible.
From the total number of organisms, `r format(n_sp_names_sp_level, big.mark = ",")` were originally identified to the species-level, `r format(n_sp_names_gen_level, big.mark = ",")` to the genus level, and no taxonomic information was avaliable for the remaining `r format(n_no_tax_info, big.mark = ",")`. 
 
We were able to confirm the validity of `r format(n_sp_correct_in_data, big.mark = ",")` of the cannonical names used to identify organisms to the species level (roughly `r scales::percent(n_sp_correct_in_data/n_sp_names_sp_level, 1)`). 
We assessed the validity of a scientific name by querying the Global Names Resolver database (https://resolver.globalnames.org) which includes taxonomic data from `r dplyr::n_distinct(taxize::gnr_datasources()$title)` taxonomic sources. 
Specifically, we used the function `gnr_resolve` from the R package `taxize `r as.character(packageVersion('taxize'))`` [@chamberlain_taxize_2013; @chamberlain_taxize_2019].

From the remaining `r format(n_sp_names_sp_level - n_sp_correct_in_data, big.mark = ",")` names we were unable to validate, we were able to identify and correct `r format(n_sp_corrected_typos, big.mark = ",")` spelling mistakes. 
These spelling mistakes were corrected automatically by fuzzy matching the canonical names in our data sources with tgose in the Global Names Resover database. 
In rare occasions, the fuzzy matching algorithm can suggest a cannonical name that has similar spelling but that correspond to an organism in a completely different taxonomic group, often in a different kingdom. 
We, therefore, checked the taxonomic hierarchy of suggested names to confirm that it corresponded to our expectations. 
We retrieved all taxonomic hierarchies from the National Center for Biotechnology Information taxonomic database (https://www.ncbi.nlm.nih.gov/taxonomy). 

Subsequently, we checked for possible synonyms of the canonnical names in our data sources. 
Using data from the Integrated Taxonomic Information System database (http://www.itis.gov), we found synonyms and alternative names for `r format(n_sp_with_synonyms, big.mark = ",")` of those in our data sources.
Finding these alternative names was important due to two reasons. 
First, because we wanted to be able to identify the cases in which the same organism might have been recorded with different cannonical names in different data sources. This can ocurr not only with synonyms but also when the cannonical name has been changed or when there are widely used ortographic variants.
Second, because retrieving occurrence data is often only possible using the latest accepted/valid name for a particular species.

All together, from the `r format(n_sp_names_sp_level - n_sp_correct_in_data, big.mark = ",")` names we were unable to validate, it was not possible to automatically correct `r format(n_sp_not_found, big.mark = ",")` of them. 
We then manually consulted multiple online databases, chiefly Wikispecies (https://species.wikimedia.org/), and looked for cannonical names that resembled the unvalidated names and matched the geographic and taxonomic spectations. 
We were able to further correct `r n_sp_manual_corrections` names in this fashion.
Most manual corrections were of species names that have been abbreviated or had more than two spelling mistakes. 
A complete list of manual name corrections can bee seen in Table \@ref(tab:tab-manual-name-corrections).

```{r}
loadd(recoded_interactions, clean_interactions)

n_sp_ids <- recoded_interactions %>%
  select(-loc_id) %>%
  purrr::map(n_distinct) %>%
  unlist() %>% sum()

n_names_in_multiple_references_after <- recoded_interactions %>%
  tidyr::gather(key = "guild", "sp_id", pla_id, ani_id) %>%
  left_join(study_per_location, by = "loc_id") %>%
  group_by(sp_id) %>%
  summarise(n_ref = n_distinct(reference)) %>%
  filter(n_ref > 1) %$%
  n_distinct(sp_id)

stats_clean_int <- clean_interactions %>%
  purrr::map(n_distinct)
```

This cleaning process allowed us to match further `r n_sp_names - n_sp_ids` names across data sources and, by doing so, identify another `r n_names_in_multiple_references_after - n_names_in_multiple_references_orig` species that were present in more than one study. 
It also allowed us to identify `r numbers2words(length(problematic_networks))` problematic networks in which some names were included as both plants and pollinators. 
These networks were removed from further analysis. 
Sometimes interaction data was recorded at multiple points in time within or across our data sources. 
When this was the case, we combined them into one single interaction network per location. 
The cleaning process resulted on a total of `r format(stats_clean_int$pla_id, big.mark = ",")` plants and `r format(stats_clean_int$ani_id, big.mark = ",")` pollinator species distributed across `r format(stats_clean_int$loc_id)` locations arround the globe (Figure \@ref(fig:fig-worldmap)).

```{r fig-worldmap, fig.width = width('single'), fig.cap = paste("Worldwide distribution of pollination communities included in this study. We compile data from", length(interaction_citations$BIBTEXKEY), "published studies", all_int_cites, "which contain interaction information between plant and their pollinators at", dplyr::n_distinct(int_metadata$loc_id), "distinct locations.")}
readd(fig_worldmap)
```

### Climatic and habitat

<!-- These data are for use in research projects only. A 'research project' is any project carried out by an individual or organized by a university, a scientific institute, or similar organization (private or public) for non-commercial research purposes only. Results based on these data must be submitted for publication in the open literature without any delay linked to commercial objectives. An additional restriction is that the UK Met. Office/Hadley Centre requires strict adherence to the conditions set forth in their "License Statement", which applies to their model output: "These data are licensed for use in Research Projects only. A 'Research Project' is any project organised by a university, a scientific institute, or similar organisation (private or public), for non-commercial research purposes only. A necessary condition of the recognition of non-commercial purposes is that all the results obtained are openly available at delivery costs only, without any delay linked to commercial objectives, and that the research itself is submitted for open publication. -->

<!-- Proper citation and acknowledgement -->
<!-- In first making reference to the data obtained from the CliMond archive, please include the phrase "the CliMond dataset (Kriticos et al. 2012)". Subsequent references within the same publication might refer to the data with terms such as " CliMond data", "the CliMond multi-model dataset", "the CliMond archive", or the " CliMond dataset". Please note the following critical references: -->

<!-- CliMond: -->
<!-- Kriticos DJ, Webber BL, Leriche A, Ota N, Bathols J, Macadam I, Scott JK (2012) CliMond: global high resolution historical and future scenario climate surfaces for bioclimatic modelling. Methods in Ecology and Evolution, 3, 53-64.  -->
<!-- WorldClim: -->
<!-- Hijmans RJ, Cameron SE, Parra JL, Jones PG, Jarvis A (2005) Very high resolution interpolated climate surfaces for global land areas. International Journal of Climatology, 25, 1965-1978.  -->
<!-- CRU: -->
<!-- Mitchell TD, Carter TR, Jones PD, Hulme M, New M (2004) A comprehensive set of climate scenarios for Europe and the globe: the observed record (1900-2000) and 16 scenarios (2000-2100).University of East Anglia, Norwich,UK, pp. 30.  -->
<!-- CMIP3: -->
<!-- Meehl, G. A., C. Covey, T. Delworth, M. Latif, B. McAvaney, J. F. B. Mitchell, R. J. Stouffer, and K. E. Taylor (2007) The WCRP CMIP3 multi-model dataset: A new era in climate change research, Bulletin of the American Meteorological Society, 88, 1383-1394.  -->
<!-- CSIRO Mk 3.0: -->
<!-- Collier MA, Dix MR, Hirst AC (2007) CSIRO Mk3 climate system model and meeting the strict IPCC AR4 data requirements. MODSIM07 International Congress on Modelling and Simulation : Land, water & environmental management : integrated systems for sustainability, 10-13 December [2007] : proceedings (eds Oxley L, Kulasiri D), Christchurch, pp. 582-588. Modelling and Simulation Society of Australia and New Zealand.  -->
<!-- Miroc-H: -->
<!-- Hasumi H, Emori S (2004) K-1 coupled model (MIROC) description, K-1 technical report, 1. Center for Climate System Research, University of Tokyo, Tokyo, 34 pp. -->


See topography data from Species distribution model transferability and model grain size – finer may not always be better

# Results

# Discussion 

# Acknowledgements

`r paste(drake::readd(acknowledgements), collapse = "")`

# References
