---
title: "Environmental stress affects niche breadth in plant-pollinator communities"
subtitle: "Supplementary information"
author: "E. Fernando Cagua, Hugo J. Marrero, Jason M. Tylianakis, Daniel B. Stouffer"
bibliography: 
  - bibliography.bib
  - int-bibliography.bib
# csl: pnas.csl
output: 
  bookdown::pdf_document2:
    # base_format: rticles::peerj_article
    keep_tex: yes
    number_sections: false
toc: false
fontsize: 11pt
# classoption: a4paper
geometry: 
  - textwidth=33em
  - textheight=48\baselineskip
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{xcolor}
  - \usepackage{setspace}
  - \usepackage{caption}
editor_options: 
  chunk_output_type: console
---

\onehalfspacing
\captionsetup[table]{skip=10pt}

<!-- To add an S before each figure and table -->
\renewcommand\thefigure{S\arabic{figure}}    
\setcounter{figure}{0}   

\renewcommand\thetable{S\arabic{table}}    
\setcounter{table}{0}   

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, 
                      warning = FALSE)
options(booktabs = TRUE, 
        knitr.kable.NA = "-")
library(dplyr)
```

# Supplementary methods

```{r scientific-names-cleaning-info}
loadd(checked_sp_names, spp, manual_name_corrections, problematic_networks)
loadd(int_metadata)

n_sp_names <- n_distinct(spp$sp_name) 

n_sp_names_subsp_level <- filter(spp, is_subspecies) %$%
  n_distinct(sp_name)

n_sp_names_sp_level_no_subsp <- filter(spp, !sp_unidentified, !is_subspecies) %$% 
  n_distinct(sp_name)

n_sp_names_sp_level <- filter(spp, !sp_unidentified) %$% 
  n_distinct(sp_name)

n_sp_names_gen_level <- filter(spp, sp_unidentified, !gen_unidentified) %$% 
  n_distinct(sp_name)

n_no_tax_info <- filter(spp, gen_unidentified) %$% n_distinct(sp_name) 

n_sp_correct_in_data <- checked_sp_names %>%
  filter(good_queried_sp_name) %$%
  n_distinct(queried_sp_name)

n_sp_corrected_typos <- checked_sp_names %>%
  filter(corrected_typo) %$% 
  n_distinct(queried_sp_name)

sp_with_synonyms <- checked_sp_names %>%
  filter(alt_sp_name_found) 

n_sp_with_synonyms <- sp_with_synonyms %$%
  n_distinct(queried_sp_name)

sp_not_found <- checked_sp_names %>%
  filter(!good_queried_sp_name) %>% 
  filter(!alt_sp_name_found, !corrected_typo) 

n_sp_not_found <- sp_not_found %$%
  n_distinct(queried_sp_name)

n_sp_manual_corrections <- manual_name_corrections %>%
  filter(!is.na(sp_name)) %$%
  n_distinct(queried_sp_name)

study_per_location <- int_metadata %>%
  distinct(loc_id, reference)

n_names_in_multiple_references_orig <- spp %>%
  left_join(study_per_location, by = "loc_id") %>%
  group_by(sp_name) %>%
  summarise(n_net = n_distinct(loc_id), 
            n_ref = n_distinct(reference)) %>%
  filter(n_ref > 1) %$%
  n_distinct(sp_name)

all_int_cites <- paste0("@", interaction_citations$BIBTEXKEY) %>%
  glue::glue_collapse(sep = "; ") %>%
  paste0("[", ., "]")
```

Data was obtained from the Web of Life database [@fortuna_web_2014] which includes data from `r length(interaction_citations$BIBTEXKEY)` published studies `r all_int_cites`.

All together, interaction data from the included studies included `r format(n_sp_names, big.mark = ",")` unique organism names. 
From these `r format(n_names_in_multiple_references_orig, big.mark = ",")` were present in more than one study.
From the total number of organisms, `r format(n_sp_names_subsp_level, big.mark = ",")` were identified to the subspecies or variety level, `r format(n_sp_names_sp_level_no_subsp, big.mark = ",")` to the species level, `r format(n_sp_names_gen_level, big.mark = ",")` to the genus level, whereas the remaining `r format(n_no_tax_info, big.mark = ",")` were unidentified. 
As the species level was the most common taxonomic rank available in our interaction datasets, in all further analysis, we grouped together subspecies or varieties within the same species. 

We were able to confirm the validity of `r format(n_sp_correct_in_data, big.mark = ",")` of the scientific names used to identify organisms (roughly `r scales::percent(n_sp_correct_in_data/n_sp_names_sp_level, 1)`). 
We assessed the validity of a name by querying the Global Names Resolver database (https://resolver.globalnames.org) which includes data from `r dplyr::n_distinct(taxize::gnr_datasources()$title)` taxonomic sources. 
We accessed this database using the function `gnr_resolve` from the R package `taxize` `r as.character(packageVersion('taxize'))` [@chamberlain_taxize_2013; @chamberlain_taxize_2019].

From the remaining `r format(n_sp_names_sp_level - n_sp_correct_in_data, big.mark = ",")` names we were unable to validate, we were able to identify and correct `r format(n_sp_corrected_typos, big.mark = ",")` that contained spelling mistakes. 
These spelling mistakes were corrected automatically by fuzzy matching the canonical names in our data sources with those in the Global Names Resolver database. 
However, on rare occasions, the fuzzy matching algorithm can suggest a scientific name that has a similar spelling, but that corresponds to an organism in a different taxonomic group, often a separate kingdom. 
To address this potential problem, we checked the taxonomic hierarchy of suggested names and confirmed that it matched our expected taxon.
We retrieved all taxonomic hierarchies from the National Center for Biotechnology Information taxonomic database (https://www.ncbi.nlm.nih.gov/taxonomy). 

As species names are constantly changing, we subsequently checked for possible synonyms of the canonical names in our data sources. 
Using data from the Integrated Taxonomic Information System database (http://www.itis.gov), we found synonyms and alternative names for `r format(n_sp_with_synonyms, big.mark = ",")` species.  
Finding these alternative names was required for two main reasons. 
First, because we wanted to be able to identify the cases in which the same species might have been recorded with different names in various data sources. 
This can occur not only when the canonical name has been changed but also when there are widely used orthographic variants.
Second, because retrieving occurrence data is often only possible using the latest accepted/valid name for a particular species.

All together, from the `r format(n_sp_names_sp_level - n_sp_correct_in_data, big.mark = ",")` names we were unable to validate, it was not possible to automatically correct or find synonyms `r format(n_sp_not_found, big.mark = ",")` of them. 
We then manually consulted multiple online databases, chiefly Wikispecies (https://species.wikimedia.org/), and looked for canonical names that both, resembled the unvalidated names and matched the geographic and taxonomic expectations. 
In this fashion, we were able to further correct `r n_sp_manual_corrections` names. 
Most manual corrections were made on names that have been abbreviated or had more than two spelling mistakes. 
A complete list of manual name corrections can bee seen in Table \@ref(tab:tab-manual-name-corrections).

```{r tab-manual-name-corrections}
loadd(spp, manual_name_corrections)
manual_name_corrections %>%
  inner_join(spp, by = c("queried_sp_name" = "sp_name")) %>%
  distinct(queried_sp_name, sp_name, guild) %>%
  filter(!is.na(sp_name)) %>%
  arrange(guild, queried_sp_name, sp_name) %>%
  mutate(sp_name = tidyr::replace_na(sp_name, "no replacement found"), 
         sp_name = kableExtra::cell_spec(sp_name, 
                                         format = "latex", 
                                         italic = TRUE), 
         guild = if_else(guild == "pla", "plant", "pollinator")) %>%
  knitr::kable(format = "latex", 
               caption = "Manually corrected canonnical names. More than one correct name have been included when an accepted/valid synonym the cannonical name exists.", 
               col.names = c("incorrect name", "corrected name", "guild"),
               booktabs = TRUE, 
               escape = F) %>%
  kableExtra::kable_styling(font_size = 8) %>%
  kableExtra::collapse_rows(1, latex_hline = "none", valign = "top")
```

```{r}
loadd(recoded_interactions, int_metadata)

n_sp_ids <- recoded_interactions %>%
  select(-loc_id) %>%
  purrr::map(n_distinct) %>%
  unlist() %>% sum()

n_names_in_multiple_references_after <- recoded_interactions %>%
  tidyr::gather(key = "guild", "sp_id", pla_id, ani_id) %>%
  left_join(study_per_location, by = "loc_id") %>%
  group_by(sp_id) %>%
  summarise(n_ref = n_distinct(reference)) %>%
  filter(n_ref > 1) %$%
  n_distinct(sp_id)


data_sources_mult_nets <- int_metadata %>%
  group_by(loc_id) %>%
  summarise(n_nets = n_distinct(net_name)) %>%
  filter(n_nets > 1) 

n_data_sources_mult_nets <- nrow(data_sources_mult_nets)
```

This cleaning process allowed us to match further `r n_sp_names - n_sp_ids` names across data sources and, by doing so, identify another `r n_names_in_multiple_references_after - n_names_in_multiple_references_orig` species that were present in more than one study. 
The process also allowed us to identify `r numbers2words(length(problematic_networks))` problematic data sources in which some names were included as both plants and pollinators. 
These data sources were removed from further analysis. 
In `r numbers2words(n_data_sources_mult_nets)` of our data sources interaction data was recorded at multiple points in time. 
When this was the case, we combined interaction data into one single interaction network. 

```{r tab-studies, eval = F}
loadd(int_metadata)
loadd(interaction_citations)

simple_metadata <- int_metadata %>%
  filter(stringr::str_detect(reference, pattern = "unpubl", negate = TRUE)) %>%
  group_by(loc_id) %>%
  summarise(location = paste0(round(mean(lon), 2), ", ", 
                              round(mean(lat), 2)), 
            name = first(loc_name), 
            ref = first(reference)) %>%
  # first word in the references is the main author (usually)
  mutate(first_author = stringr::str_extract(ref, "^\\w+"), 
         # the first number of four digits is the year
         year = stringr::str_extract(ref, "[0-9]{4}"), 
         key = paste(tolower(first_author), year, sep = "_")) %>%
  select(-first_author, -year)


interaction_citations %>%
  mutate(first_author = stringr::str_extract(BIBTEXKEY, "[^_-]+"), 
         year = stringr::str_extract(BIBTEXKEY, "[0-9]{4}"), 
         key = paste(tolower(first_author), year, sep = "_"), 
         bibtex_key = BIBTEXKEY) %>%
  select(bibtex_key, key) %>%
  fuzzyjoin::stringdist_full_join(simple_metadata, 
                                  by = c(key = "key"), 
                                  max_dist = 6, 
                                  distance_col = "dist") %>% 
  group_by(key.y) %>%
  filter(dist == min(dist)) %>%
  ungroup() %>%
  select(location, name, bibtex_key) %>%
  mutate(bibtex_key = paste0("@", bibtex_key, "")) %>%
  knitr::kable(format = "markdown", 
               escape = F, 
               booktabs = T, 
               col.names = c("coordinates", "site", "reference"),
               caption = "Sites included in this study")
  kableExtra::kable_styling(font_size = 8)
``` 

```{r eval = FALSE, tab-sp-synonym-reasons}
sp_with_synonyms %>%
  filter(!is.na(itis_reason)) %>%
  group_by(queried_sp_name) %>%
  summarise(itis_reason = first(itis_reason)) %>%
  group_by(itis_reason) %>%
  count(sort = T) %>%
  knitr::kable("latex", 
               col.names = c("reason of species alternative name", "n"), 
               caption = "tablesita", 
               booktabs = TRUE) %>%
  kableExtra::kable_styling(font_size = 8) %>%
  kableExtra::collapse_rows(columns = 1)
```

# Supplementary references
